---
title: "Tables"
author: "Drew Clayson"
format: html
editor: visual
---

## Data Gathering

I am doing my data based on the G20 members to keep a concise table. I filter the dates down to ensure that there are no missing values (some data have not been updated since 2015).

```{r, results = FALSE}
library(wbstats)
library(tidyverse)
library(knitr)
library(kableExtra)

raw_data <- wb_data(country = c("ARG", "AUS", "BRA", "CAN", "CHN", "FRA", "DEU", "IND", "IDN", "ITA", "KOR", "JPN", "MEX", "RUS", "SAU", "ZAF", "TUR", "GBR", "USA", "EU"), indicator = c("NY.GDP.MKTP.KD", "NY.GDP.PCAP.KD.ZG", "EN.ATM.CO2E.KT", "EG.ELC.RNEW.ZS", "NY.GNS.ICTR.ZS"))
clean_data <- raw_data %>% filter(date < 2016) %>% filter(date > 1999) %>% select(c(3:9))
colnames(clean_data) <- c("Country", "Date", "Percent Renewable", "CO2", "GDP(billions)", "GDP Per Capita Growth", "Savings Rate")
clean_data$`GDP(billions)` <- clean_data$`GDP(billions)`/1000000000
```

For my first table, I will be making a relatively simple table with a few added flairs pointing out interesting information. I understand that it might be slightly redundant to simply highlight negative numbers in red, however I think it makes it much easier for readers to glean simple information quickly. Furthermore I chose to highlight relatively large savings rates in green and found that they tend to all be from Asia, which is perhaps an interesting insight!

```{r}
clean_data$`GDP Per Capita Growth` <- ifelse(
  clean_data$`GDP Per Capita Growth` < 0,
  cell_spec(round(clean_data$`GDP Per Capita Growth`,digits = 2), color = "red", bold = TRUE),
  cell_spec(round(clean_data$`GDP Per Capita Growth`,digits = 2))
)
clean_data$`Savings Rate` <- ifelse(
  clean_data$`Savings Rate`>30,
  cell_spec(round(clean_data$`Savings Rate`,digits = 2), color = "green"),
  cell_spec(round(clean_data$`Savings Rate`, digits = 2))
)

tbl <- clean_data %>%
  ungroup() %>%
  filter(Date == 2015) %>%
  arrange(desc(`GDP(billions)`)) %>%
  select(Country, `GDP(billions)`, `GDP Per Capita Growth`, `Savings Rate`, CO2, `Percent Renewable`) %>%
  kbl(escape = FALSE,
    digits = 2) %>%
  add_header_above(c(" ", "Economic Data " = 3, "Environmental Data" = 2))
tbl %>%
  kable_material(lightable_options = c("striped", "hover", "")) %>%
  htmltools::HTML() %>%
  shiny::div() %>%
  sparkline::spk_add_deps()
```

From here, I will try to make use of the time series data using sparkline. I would really only like to include time series of GDP, CO2, and Percent Renewable. I could do time series of the other data, however I don't believe it to be particularly interesting. The chart will still use the same data as above so all color highlights remain the same, which is nice functionality.

```{r}
library(sparkline)
# Create similar table to the one above,  however mutating to add 3 more variables
tbl2 <- clean_data %>%
  filter(Date == 2015) %>%
  arrange(desc(`GDP(billions)`)) %>%
  select(Country, `GDP(billions)`, `GDP Per Capita Growth`, `Savings Rate`, CO2, `Percent Renewable`) %>%
  mutate(plt_GDP = NA, plt_CO2 = NA, plt_PercentRenewable = NA) %>% as.data.frame()
# Generate plots using a for loop
for(i in tbl2$Country){
  # Data for GDP
  country_GDP <- clean_data %>%
    filter(Country==i) %>%
    arrange(Date) %>%
    pull(`GDP(billions)`)
  tbl2[which(tbl2$Country == i), "plt_GDP"] <- spk_chr(country_GDP)
  # Data for CO2
  country_CO2 <- clean_data %>%
    filter(Country==i) %>%
    arrange(Date) %>%
    pull(CO2)
  tbl2[which(tbl2$Country == i), "plt_CO2"] <- spk_chr(country_CO2)
  # Data for Renewables
  country_PercentRenewable <- clean_data %>%
    filter(Country==i) %>%
    arrange(Date) %>%
    pull(`Percent Renewable`)
  tbl2[which(tbl2$Country == i), "plt_PercentRenewable"] <- spk_chr(country_PercentRenewable)
}
tbl2_full <- tbl2 %>%
  select(Country, `GDP(billions)`, plt_GDP, `GDP Per Capita Growth`, `Savings Rate`, CO2, plt_CO2, `Percent Renewable`, plt_PercentRenewable) %>%
  kbl(escape = FALSE,
    col.names = c("Country",
                    "GDP (billions)",
                    "GDP Trend",
                    "GDP per Capita Growth",
                    "Savings Rate",
                    "CO2 Emissions",
                    "Emission Trend",
                    "Percent Renewable",
                    "Renewability Trend"
                    ),
      digits = 2) %>%
  add_header_above(c(" ", "GDP" = 3, " ", "Environmentals" = 4)) %>%
  kable_styling(font_size=28, fixed_thead = TRUE)
tbl2_full %>%
  htmltools::HTML() %>%
  shiny::div() %>%
  sparkline::spk_add_deps()
```
